---
title: "Final Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(skimr)
library(broom)
library(olsrr)
library(viridis)
library(scales)
library(readr)
library(stringr)
library(car)
 fastfood_calories <- read_csv("fastfood_calories.csv")
```

# Analyzing Fast-food nutrients

The data seems to be in a fine format. The "salad" column does not seem to provide any information so it is better to drop it. Another notable shortcoming of the dataset is that it does not include the weight of the meals which would allow us for a better analysis. Keeping that in mind the main question is: if you are hungry and you really want to go to a fast food restaurant, which one should you choose based on its overall nutrient values?
To allow a better understanding it is better to change the raw values into percentage values. We use the FDA recommendations for a 2000 calorie diet to do that. (https://www.fda.gov/media/135301/download) It is clear from the source of original dataset (https://fastfoodnutrition.org/mcdonalds/artisan-grilled-chicken-sandwich) that it uses the same measurements (g,mg, etc.) as the FDA recommendations. Vitamins only have to be converted by decimals as they were in percent measurement from the start. 

```{r, message = FALSE, include = FALSE}


str(fastfood_calories)
fastfood_calories <- fastfood_calories %>%
  select(1:17) %>%
  mutate(total_fat_perc = total_fat/78)%>%
  mutate(sat_fat_perc = sat_fat/20)%>%
  mutate(trans_fat_perc = trans_fat/20)%>%
  mutate(cholesterol_perc = cholesterol/300)%>%
  mutate(sodium_perc = sodium/2300)%>%
  mutate(total_carb_perc = total_carb/275)%>%
  mutate(fiber_perc = fiber/28)%>%
  mutate(sugar_perc = sugar/50)%>%
  mutate(protein_perc = protein/50)%>%
  mutate(vit_a_perc = vit_a/100)%>%
mutate(vit_c_perc = vit_c/100)%>%
mutate(calcium_perc = calcium/100)%>%
  mutate(calories_perc = calories/2000)
```

## Creating percentage table
This table shows the nutrients of the items as a percentage of RDI.
```{r, include = TRUE, message = FALSE }
fast_food_percent <- fastfood_calories %>%
    select(1:4, 18:30)
```
## Which places to go?

Which restaurant offers food with the least overall calories?

```{r}
fast_food_percent %>%
      group_by(restaurant)%>%
     summarize(mean_calories = mean(calories))%>%
     arrange(mean_calories)

```

### Creating information about the most unhealthy foodstuffs

It would be good to see which are the articles in the list that have a value of more than 1 (to see which article will cover more than what you need for one day and thus is probably best avoided)
```{r pressure, echo=FALSE}
avoid_food <- fast_food_percent %>%
     select(2:3,5:17)
avoid_food %>%
     filter_at(.vars= 3:15, .vars_predicate = any_vars(. > 1)) %>%
     select(1:2)
```
It would also be good to know what are the items that have high and low salt and fat (the two most common ingredients, fast food chaines are criticized for). A plot is made to visualize the items that have very high salt and/or fat in them and which restaurant produces them and also a table is made to see which are the exact products that are most dangerous.



```{r, include = TRUE, message = FALSE, echo = FALSE }
ggplot(fast_food_percent, aes(x = total_fat_perc, y  = sodium_perc, color = restaurant)) + geom_point() + scale_x_continuous(name = "Total Fat %", limits = c(0,1.1), labels = percent) + scale_y_continuous(name = "Salt %", limits = c(0,2), labels = percent) + ggtitle("Fat and salt in fast food articles expressed by percentage of RDI")      


low_fat_salt <- fast_food_percent %>%
  filter(sodium_perc <= 0.25, total_fat_perc <= 0.15)

 high_fat_salt <- fast_food_percent %>%
     filter(sodium_perc >= 1 | total_fat_perc >= 1)
 low_fat_salt <- low_fat_salt %>%
          mutate(healthy = ifelse(.>1, 1,0))
 high_fat_salt <- high_fat_salt  %>%
          mutate(healthy = ifelse(.>1, 2,0))
 print(high_fat_salt)
 
```
It seems that some of the items are very high in salt and even though you can find some items that will cover one's fat requirements for a day there is no item that exceeds the 100% limit by too much.
We can also see that there are quite a few items that are dangerously high in both fat and salt. It would be nice to know more about them.

On the other hand there are some that have very low fat and salt.
It would be good to see whether these are in fact healthy picks in a restaurant.
Which nutrients do these food consist of? 

To explore the low end of this scale, we need a (somewhat arbitrary) cut-off point in the data. For this analysis I will use 15 and 25% percent points which seems like a good limit as the values are quite small but there are still a fair amount of points inside. The points need to be lower in both axis to qualify.
For the high end the points the limits are less arbitrary. I'm interested in the items that reach 100% in at least one of the categories, so it will be enough for a point to be larger than 100% on one of the scales.
The rest of the distribution is the Normal category.




```{r, include = TRUE, message = FALSE, echo = FALSE}
fast_food_percent <- fast_food_percent %>%
     mutate(healthy = if_else(total_fat_perc >=1 | sodium_perc >=1, "High", if_else(total_fat_perc <= 0.15 & sodium_perc <= 0.25, "Low", "Normal")))

fast_food_percent %>%
     group_by(healthy)%>%
    summarise(mean_calories = mean(calories)) %>%
        arrange(desc(mean_calories))
fast_food_percent <- fast_food_percent %>%
     mutate(vitamins = (vit_a_perc + vit_c_perc + calcium_perc)/3)

ggplot(fast_food_percent, aes(x = vitamins, y = fiber_perc, color = healthy)) + geom_point() + geom_jitter(height = 0.1, width  = 0.1) + scale_color_viridis_d() + ggtitle("Distribution of High and Low category food by fiber and vitamin contained ")

ggplot(fast_food_percent, aes(x = restaurant)) + geom_bar(aes(fill=healthy)) + ggtitle("Count of items by restaurant and colored by the ratio of their quality of food") + scale_fill_viridis_d()

 
```


It is clear that Low category food contains much less calorie than High Category food but the scatterplot also makes it clear that most of the Low salt and fat food are also low in vitamins and fiber while the high category food are higher in both of those. 
There are exceptions though, some red dots are quite low on both axis while some green go above average.
Again, portion size is missing which might be an explanation for this phenomenon.


## Searching for healthy food
Unhealthy food is already listed, now it is time to search for healthy food that have high vitamins and high fibre but low fat and low sodium. Also it is time to add sugar to the unhealthy spectrum and see what are the items that are low on that as well.

```{r, echo = FALSE}
fast_food_percent <- fast_food_percent %>%
     mutate(recommended = ifelse(healthy == "Low" & sodium_perc <= 0.25 & total_fat_perc <= 0.15 & sugar_perc <= 0.25, 1,0))
low_all_food <- fast_food_percent %>%
  filter(recommended == 1) %>%
  arrange(desc(calories))

ggplot(low_all_food, aes(x = restaurant)) + geom_bar() +ggtitle("Number of low sugar, fat and salt items by restaurant ")

ggplot(low_all_food, aes(x = calories, y = vitamins, color = restaurant)) + geom_point() + geom_text(aes(label= ifelse(calories>=200 & vitamins >= 0.15, item, ""))) + ggtitle("Low salt and fat food displayed by calorie and vitamin", subtitle = "Displayed names are for food over 200 calories and over 15% in vitamins C,A and Calcium")

low_all_food %>%
       filter(calories >= 200) %>%
     arrange(desc(calories))%>%
     select(2:3)

```

Subway and Taco Bell has most of this kind of food but the one that will cover most of your vitamin needs for not too many claories comes from McDonalds.
A list of all recommended food containing at least 200 calories to lower hunger but containing low salt, sugar and fat is also provided (as the scatterplot missed many data-points due to NAs on vitamins.)

## Testing
Note: I found it hard to come up with a test that is interesting in any ways as all the variables relate to the ingredients of the food. For instance a variable containing ratings of the items would have been very useful. 


What are the ingredients that increase calories the most?

First it is time to replace NAs in the fiber and protein variable with the mean.
```{r, echo = FALSE}
just_fiber <- fastfood_calories%>%
         select(12)
just_fiber <- just_fiber %>%
     drop_na()
just_fiber%>%
    summarise(mean_fiber = mean(fiber))

just_protein <- fastfood_calories%>%
         select(14)
just_protein <- just_protein%>%
     drop_na()
just_protein%>%
    summarise(mean_protein = mean(protein))
 fastfood_calories <- fastfood_calories%>%
     mutate(fiber = replace_na(fiber,4.1))
 
 fastfood_calories <- fastfood_calories%>%
     mutate(protein = replace_na(protein,27.9))

```



The prerequisites for linear regression seem to be met. VIF scores are below 5 and the distribution seems to be normal and homogeneous (with some deviation at the high end).
There is one very influential outlier in the datasaet, Sonic's ultimate chicken club (article 128) 
Smokehouse Beef Short Rib Sandwich (172) and Burger King's Bacon King (197) are not outliers based on Cook's distance but they stand out from the otherwise good distribution. 
After checking these datapoints it seems that there is no mechanical reason to exclude these points from the analysis. Their values do not seem to be typos, out of range or sampling errors so I will not remove them from the dataset but will try a second analysis which might yield a better result for predicting calories.


```{r echo = FALSE}
first_model<- lm(calories~total_fat + sodium + total_carb + fiber + sugar + protein, data = fastfood_calories)
first_model %>%
    augment() %>%
    arrange(desc(.cooksd)) %>%
    top_n(15)
plot(first_model, 1)
plot(first_model, 2)
plot(first_model, 3)

input_table <- fastfood_calories%>%
          select(6,10:14)
cor(input_table)
vif(first_model)
```
```{r}
summary(first_model)

second_model <- lm(calories ~ total_carb + total_fat + protein, data = fastfood_calories)
summary(second_model)
```
The first model has an R2 of 97% which is very high. The result is not surprising at all. Calories depend very much on the nutrients of the food. 
A second model is considered where only the significant predictors are kept. The R2 doesn't change indicating that the other factors do not play a significant role in the calories of fast food.

## Duel of McDonald's and Burger King

Two of the most famous chains are Burger King and McDonald's. I would like to check if there is a difference in the calorie of their items.
To compare the categories will first be converted to a numeric value (1 for McDonald's and 2 for Burger King).

```{r message = FALSE, echo = FALSE}
Mc_table <- fastfood_calories %>%
    filter(restaurant == "Mcdonalds")
 BK_table <-  fastfood_calories %>%
         filter(restaurant == "Burger King")
 duel_table <- full_join(Mc_table, BK_table)
 
  duel_table <- duel_table%>% 
     mutate(restaurant = str_replace_all(restaurant, "Mcdonalds", "1"))
 duel_table <- duel_table%>% 
         mutate(restaurant = str_replace_all(restaurant, "Burger King", "2"))
  duel_table <- duel_table%>% 
        mutate(restaurant = as.factor(restaurant))
  
  duel_model <- lm(calories~restaurant, data = duel_table)
  summary(duel_model)
  
  ggplot(duel_table, aes(x = restaurant, y = calories, color = restaurant)) + geom_boxplot( show.legend = FALSE) + ggtitle("Avarage calories for McDonalds and Burger King") + scale_x_discrete(name = c("Restaurant") ,  labels=c("McDonalds","Burger King"))
               
```

According to the plot and to the regression there is no difference between the two restaurants' items with regards to calorie.
